---
title: "Sit2Stand Data Cleaning"
author: "Melissa Boswell"
date: "July 27, 2022"
output:
  html_document: default
  pdf_document: default
---

# Load libraries and requirements
```{r, echo = FALSE}
library(plyr)
library(readxl)
library(psych)
library(tidyverse)
library(sjmisc)
library(stddiff)
```

# Load survey data and combine with video data
```{r, echo = FALSE}
# Load Data
dataMovement <- read.csv("dataMovement.csv") # Sit-to-stand parameters
dataSurvey <- read.csv("dataSurvey.csv") # Numeric survey data
dataSurvey_text <- read.csv("dataSurvey_text.csv") # Qualitative survey data

# Combine data frames based on subject ID
colnames(dataSurvey)[which(names(dataSurvey) == "video")] <- "subjectid"
df <- join(dataMovement, dataSurvey, by="subjectid", type="inner")
colnames(dataSurvey_text)[which(names(dataSurvey_text) == "video")] <- "subjectid"
df_text <- join(dataMovement, dataSurvey_text, by="subjectid", type="inner")

# Remove participants with with bad videos
dataBad <- read_excel("REMOVE_videos.xlsx")
df <- df[! df$subjectid %in% dataBad$video,]
df_text <- df_text[! df_text$subjectid %in% dataBad$video,]

colnames(dataBad)[which(names(dataBad) == "video")] <- "subjectid"
df_removed <- join(dataBad, dataSurvey, by="subjectid", type="inner")
df_removed_text <- join(dataBad, dataSurvey_text, by="subjectid", type="inner")

df_all = rbind(df[which(colnames(df)=="Age"):which(colnames(df)=="feedback_open")], df_removed[which(colnames(df_removed)=="Age"):which(colnames(df_removed)=="feedback_open")])
df_all_text = rbind(df_text[which(colnames(df)=="Age"):which(colnames(df)=="feedback_open")], df_removed[which(colnames(df_removed)=="Age"):which(colnames(df_removed)=="feedback_open")])

#Make Numeric
df[] <- sapply(df, as.numeric)
df_all[] <- sapply(df_all, as.numeric)
df_removed[] <- sapply(df_removed, as.numeric)

#Make NA's Zero
df[is.na(df)] <- 0
df_all[is.na(df_all)] <- 0
df_removed[is.na(df_removed)] <- 0
#df_removed[is.na(df_removed)] <- 0
```

# Adjust time based on n cycles detected
```{r}
df$time <- df$time/df$n * 5
df <- subset( df, select = -n )
```

# Fix typos/units
```{r, echo = FALSE}
df$Weight[which(df_text$subjectid == "KHDHZ4c5")] <- 232 # typo
df$Weight[which(df_text$subjectid == "xVeHS4j1")] <- 163 #kg to lbs
df$Weight[which(df_text$subjectid == "hfnQyNnP")] <- 95 #kg to lbs

df_all$Weight[which(df_text$subjectid == "KHDHZ4c5")] <- 232 # typo
df_all$Weight[which(df_text$subjectid == "xVeHS4j1")] <- 163 #kg to lbs
df_all$Weight[which(df_text$subjectid == "hfnQyNnP")] <- 95 #kg to lbs
```

# Survey measure calculations
```{r, echo = FALSE}
# Survey measure calculations

# BMI
df$BMI <- df$Weight / (df$Height)^2 * 703
df_all$BMI <- df_all$Weight / (df_all$Height)^2 * 703
df_removed$BMI <- df_removed$Weight / (df_removed$Height)^2 * 703

# Promis Global Scale
# Note: These were already reverse scored in qualtrics

# Global Total
df$GT <- rowSums(df[,c("Global01", "Global02", "Global03", "Global04", "Global05", "Global06", "Global07", "Global08", "Global09", "Global10")])

# Global Mental Health T-score
df$GMH_raw <- rowSums(df[,c("Global02", "Global04", "Global05", "Global10")])
GMH_TScore <- read_excel("GMH_T-Score.xlsx")
df$GMH_TScore <- GMH_TScore$`T-Score`[match(unlist(df$GMH_raw), GMH_TScore$`Raw Score`)]

# Global Physical Health T-score
df$GPH_raw <- rowSums(df[,c("Global03", "Global06", "Global07", "Global08")])
GPH_TScore <- read_excel("GPH_T-Score.xlsx")
df$GPH_TScore <- GPH_TScore$`T-Score`[match(unlist(df$GPH_raw), GPH_TScore$`Raw Score`)]

# Physical Activity (IPAQ-SF)

# Walking to minutes accounting for people that put in minutes
df$IPAQ_SF_6_1[which(df$IPAQ_SF_6_1 <= 10)] <- df$IPAQ_SF_6_1[which(df$IPAQ_SF_6_1 <= 10)] * 60

# Remove if < 10 minutes
df$IPAQ_SF_2_2[which(df$IPAQ_SF_2_2 < 10)] = 0
df$IPAQ_SF_4_2[which(df$IPAQ_SF_4_2 < 10)] = 0
df$IPAQ_SF_6_1[which(df$IPAQ_SF_6_1 < 10)] = 0

# New variables
df$vigMins <- df$IPAQ_SF_2_2
df$modMins <- df$IPAQ_SF_4_2
df$walkMins <- df$IPAQ_SF_6_1

# Truncate vigMins, modMins, and walkMins if >180 mintes
df$vigMins[which(df$vigMins > 180)] = 180
df$modMins[which(df$modMins > 180)] = 180
df$walkMins[which(df$walkMins > 180)] = 180

# Total time per week
df$vigTime <- df$IPAQ_SF_1_1 * df$vigMins
df$modTime <- df$IPAQ_SF_3_1 * df$modMins
df$walkTime <- df$IPAQ_SF_5_1 * df$walkMins
df$totalTime <- rowSums(df[,c("vigTime", "modTime", "walkTime")])

# Remove outliers
# Not currently removing outliers because there are a lot over the outlier time; likely because many people are active students in California
# df <- df[-which(df$totalTime > 960)]

# Calculate metabolic minutes
df$vigMET <- 8 * df$vigTime
df$modMET <- 4 * df$modTime
df$walkMET <- 3.3 * df$walkTime
df$totalMET <- rowSums(df[,c("vigMET", "modMET", "walkMET")])

df$IPAQ_days <- df$IPAQ_SF_1_1_TEXT + df$IPAQ_SF_3_1_TEXT + df$IPAQ_SF_5_1_TEXT

# Physical activity categories
for (i in 1:length(df$totalMET)) {
if(df$totalMET[i] >= 1500){
  df$PA_cat[i] = 'High'
} else if(df$totalMET[i] <= 1500 & df$totalMET[i] >= 600){
  df$PA_cat[i] = 'Moderate'
} else {
  df$PA_cat[i] = 'Low'
}}

# Process Mindset
df$avgPEM <- rowMeans(df[,c("PES_1", "PES_2", "PES_3", "PES_4", "PES_5", "PES_6", "PES_7")])

# Adequacy Mindset
df$avgAAM <- rowMeans(df[,c("amm_1", "amm_2", "amm_3", "amm_4", "amm_5")])

# Self-Efficacy for Exercise
# Sum the score (maximum 90)
df$totalSEE <- rowMeans(df[,c("SEE_1", "SEE_2", "SEE_3", "SEE_4", "SEE_5", "SEE_6", "SEE_7", "SEE_8", "SEE_9")])

# Set NA for 0's in step count and active time
df$stepCount[df$stepCount == 0] <- NA
df$activeTime[df$activeTime == 0] <- NA

# WOMAC
df$WOMAC_pain <- df$WOMAC_1_1 + df$WOMAC_1_2 + df$WOMAC_1_3 + df$WOMAC_1_4 + df$WOMAC_1_5 

df$WOMAC_func <- df$WOMAC_3_1 + df$WOMAC_3_2 + df$WOMAC_3_3 + df$WOMAC_3_4 + df$WOMAC_3_5 + df$WOMAC_3_6 + df$WOMAC_3_7 + df$WOMAC_3_8 + df$WOMAC_3_9 + df$WOMAC_3_10 + df$WOMAC_3_11 + df$WOMAC_3_12 + df$WOMAC_3_13 + df$WOMAC_3_14 + df$WOMAC_3_15 + df$WOMAC_3_16

df$WOMAC_stiff <- df$WOMAC_2_1 + df$WOMAC_2_2

df$totalWOMAC <- df$WOMAC_pain + df$WOMAC_stiff + df$WOMAC_func

df$totalWOMAC[df$totalWOMAC == 0] <- NA
df$WOMAC_pain[df$WOMAC_pain == 0] <- NA
df$WOMAC_func[df$WOMAC_func == 0] <- NA
df$WOMAC_stiff[df$WOMAC_stiff == 0] <- NA

# Number of falls not due to sports
for (i in 1:length(df$Fall_YN)) {
if(df$Fall_YN[i] == 1 && df$Fall_Reason[i] > 1){
  df$falls[i] = df$Fall_.[i]
} else {
  df$falls[i] = 0
}}

# Binary had a fall not due to sports
for (i in 1:length(df$Fall_YN)) {
if(df$Fall_YN[i] == 1 && df$Fall_Reason[i] > 1){
  df$fallsBin[i] = 1
} else {
  df$fallsBin[i] = 0
}}

# Number of medical conditions
for (i in 1:length(df_text$MedicalConditions)) {
if(str_count(df_text$MedicalConditions[i]) == 0){
  df$numMedCond[i] = 0
} else {
  df$numMedCond[i] = str_count(df_text$MedicalConditions[i], ',') + 1
}}

# Bin number of medical conditions
for (i in 1:length(df$numMedCond)) {
if(df$numMedCond[i] == 0){
  df$numMed_1[i] = 0
} else {
  df$numMed_1[i] = 1
}}

for (i in 1:length(df$numMedCond)) {
if(df$numMedCond[i] <= 1){
  df$numMed_2[i] = 0
} else {
  df$numMed_2[i] = 1
}}

# Recode ethnicities
## White = 1 / Non-white = 0
for (i in 1:length(df_text$Ethnicity)) {
if(str_contains(df_text$Ethnicity[i], 'White')){
  df$EthBin[i] = 1
} else {
  df$EthBin[i] = 0
}}

# Recode education
## College degree or higher = 1
for (i in 1:length(df$Education)) {
if(df$Education[i] <= 3){
  df$EduBin[i] = 0
} else {
  df$EduBin[i] = 1
}}

## Separate category per race with mixed as own category
for (i in 1:length(df_text$Ethnicity)) {
if(str_contains(df_text$Ethnicity[i], ',')){
  df$EthCat[i] = 7
} else if(str_contains(df_text$Ethnicity[i], 'White')){
  df$EthCat[i] = 1
} else if(str_contains(df_text$Ethnicity[i], 'Asian')){
  df$EthCat[i] = 2
} else if(str_contains(df_text$Ethnicity[i], 'African American')){
  df$EthCat[i] = 3
} else if(str_contains(df_text$Ethnicity[i], 'Hispanic')){
  df$EthCat[i] = 4
} else if(str_contains(df_text$Ethnicity[i], 'American Indian')){
  df$EthCat[i] = 5
} else if(str_contains(df_text$Ethnicity[i], 'Other')){
  df$EthCat[i] = 6
} else if(str_contains(df_text$Ethnicity[i], 'Not')){
  df$EthCat[i] = 8
} else {
  df$EthCat[i] = 0
}}

df$backpain <- as.numeric(str_detect(df_text$MedicalConditions, "Back pain"))

df$depression <- as.numeric(str_detect(df_text$MedicalConditions, "Depression"))

df$highBP <- as.numeric(str_detect(df_text$MedicalConditions, "High blood pressure"))

df$WOMAC_total <- sum(df$WOMAC_pain + df$WOMAC_stiff + df$WOMAC_func)
```

# Variable name legend
```{r}
df_corr <- df[, c("Age", "Height", "Weight", "BMI", "Sex", "GT", "GPH_TScore", "GMH_TScore", "totalMET", "avgPEM", "avgAAM", "falling_1", "falling_2", "totalSEE", "OA_check", "totalWOMAC", "Fall_YN", "falls", "time", "right_ankle_max","right_ankle_min", "right_knee_max","right_knee_min", "right_hip_max","right_hip_min", "trunk_lean_max", "trunk_lean_ang_acc", "pelvic_max_y_acc")]
print("Variable name legend")
print("GT: Global Health total score")
print("GPH_TScore: Global physical health t-score")
print("GMH_TScore: Global mental health t-score")
print("totalMET: Physical activity (metabolic equivalent")
print("avgPEM: Process of exercise mindset")
print("avgPEM: Process of exercise mindset")
print("avgAAM: Adequacy of activity mindset")
print("falling_1: Fear of falling (0-10)")
print("OA_check: Whether the individual has osteoarthritis (1) or not (0)")
print("FAll_YN: Has had a fall in the past 12 months")
print("falls: Number of falls in the past 12 months not due to sports")
```

# View distributions
```{r}
hist(df$time, breaks=50, main="Time dist with all")
hist(df$trunk_lean_max_mean_sit2stand, breaks=50, main="Trunk lean with all")
hist(df$trunk_lean_max_ang_acc_sit2stand, breaks=50, main="Trunk lean acc with all")
```
# Bin based on age
```{r}
# set up cut-off values 
breaks <- c(18,20,30,40,50,60,70,80,100)
# specify interval/bin labels
tags <- c("18-20)","[20-30)", "[30-40)", "[40-50)", "[50-60)", "[60-70)","[70-80)", "[80+)")
# bucketing values into bins
df$ageBin <- cut(df$Age, 
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)
# inspect bins
summary(df$ageBin)

# set up cut-off values 
breaks <- c(18,30,50,70,100)
# specify interval/bin labels
tags <- c("[18-30)", "[30-50)", "[50-70)", "[70+)")
# bucketing values into bins
df$ageBin <- cut(df$Age, 
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)
# inspect bins
summary(df$ageBin)

# Bin for 50 and over
# set up cut-off values 
breaks <- c(18,50,100)
# specify interval/bin labels
tags <- c("18-50)","[50-80+")
# bucketing values into bins
df$ageGroup <- cut(df$Age, 
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)

df_text$ageGroup <- cut(df$Age, 
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)
# inspect bins
summary(df$ageGroup)
```

# Filter age groups
```{r}
df_age50plus <- filter(df, ageGroup == "[50-80+")
df_age50plus_text <- filter(df_text, ageGroup == "[50-80+")
```

# Descriptive tables
```{r}
# Removed subjects
#df_removed <- df_all[! df_all_text$subjectid %in% df_text$subjectid,]

# Simple data frames
df.descr <- df[, c("Sex", "Age", "BMI")]
df_all.descr <- df_all[, c("Sex", "Age", "BMI")]
df_removed.descr <- df_removed[, c("Sex", "Age", "BMI")]

describe(df.descr)
describe(df_all.descr)
describe(df_removed.descr)
```


# Standardized Mean Difference (between included and excluded)
```{r}
df_removed.descr$group = 0
df.descr$group = 1
dfSMD = rbind(df_removed.descr, df.descr)
stddiff.numeric(data=dfSMD, gcol=4, vcol=(1:3))
```

# 5STS Measures Descriptive tables
```{r}
# Removed subjects
#df_removed <- df_all[! df_all_text$subjectid %in% df_text$subjectid,]

# Simple data frames
df.5STS <- df[, c("time", "trunk_lean_max_mean_sit2stand", "trunk_lean_max_ang_acc_sit2stand")]
df_50.5STS <- df_age50plus[, c("time", "trunk_lean_max_mean_sit2stand", "trunk_lean_max_ang_acc_sit2stand")]

describe(df.5STS)
describe(df_50.5STS)
```

# Summarize Participant Characteristics and Variables
```{r}
# Create an extended statistics table
d.summary.extended <- df %>%
    select(Age, Height, Weight, BMI, Sex, Sex, Smoker, GT, GPH_TScore, GMH_TScore, totalMET,stepCount, activeTime, avgPEM, avgAAM, Fall_YN, falls, falling_1, falling_2, falling_3, totalSEE, OA_check, OA_yrsdiag, totalWOMAC, time, trunk_lean_max_mean_sit2stand, trunk_lean_max_ang_acc_sit2stand) %>%
    psych::describe(quant=c(.25,.75)) %>%
    as_tibble(rownames="rowname")  %>%
    print()

d.summary.extended <- df_age50plus %>%
    select(Age, Height, Weight, BMI, Sex, Sex, Smoker, GT, GPH_TScore, GMH_TScore, totalMET,stepCount, activeTime, avgPEM, avgAAM, Fall_YN, falls, falling_1, falling_2, falling_3, totalSEE, OA_check, OA_yrsdiag, totalWOMAC, time, trunk_lean_max_mean_sit2stand, trunk_lean_max_ang_acc_sit2stand) %>%
    psych::describe(quant=c(.25,.75)) %>%
    as_tibble(rownames="rowname")  %>%
    print()
```

# Participant feedback
```{r, echo = FALSE}
# Mean and SD
mean(df_all$difficult_scale)
sd(df$difficult_scale)

# Qualitative feedback
feedback <- df_all_text[, c("difficult_open", "feedback_open")]
write.csv(feedback,"feedback.csv", row.names = FALSE)
```

# Export cleaned data frame
```{r}
write.csv(df,"dataClean.csv", row.names = FALSE)
write.csv(df_text,"dataClean_text.csv", row.names = FALSE)
```
